---
title: "Learning RMarkdown"
author: "Leo"
date: "September 9, 2022"
output:
    html_document:
        toc: true
        toc_float:
            collapsed: false
            smooth_scroll: true
        number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, fig.align='center', cache=TRUE)

```


# First Section

We start a section with a single # (this is not a comment anymore). A single line of text is treated as one paragraph. 

You make a new a paragraph by leaving a blank line between lines of text. This automatically makes a new paragraph. 

In the RStudio IDE, you can spell check by hitting F7. 

# Subsections

Here we will see how to make subsections

## First Subsection

Two hashtags creates a subsection.

## Another Subsection

We just created a new subsection!

# Lists

There are two main types:

## Unordered lists

- bullet
- bullet
- bullet
    - indent
    
## Ordered Lists

1. New York
    a. coal
    a. gas
    a. wood
1. New Jersey
    1. gas
    1. wood
1. Naples
1. Detroit
1. Philadelphia
1. New Haven

# Formatting 

A _single underscore around_ text makes italics. __bold text__ looks like this with two underscore.

We use `backticks` to create `code looking text`.

# Links

[text](link)

# Using R Code!!!

Create chunk with alt+crl+i

```{r simple-addition}
#This is a comment inside a code chunk.
1+1
```
```{r make-x}
x <- 3
```
```{r multiply-x}
x * 3
```
```{r ,fig.align='left'}
y <- rnorm(200) # generates 200 random numbers
y
hist(y)
```
I want to make a hist with output not code. 
```{r ,echo=FALSE, fig.cap='A hist'}
hist(y)

```

```{r , error=TRUE}
z*2 #If one question is throwing an error, just put errow=TRUE in the code chunk so you can get partial credit on the homework. 
```

